\subsection {Session 7, Exercise 6}

The 3 theorems used in this exercise are all studied in \textbf{Introduction to the Theory of Computing 1/2}.

\lineparagraph {Exercise}

Prove that the following languages belong to coNP.
\begin{enumerate}[a)]
    \item Language of those bipartite graphs that have complete (perfect) matching.
    \item Language of those garphs that have complete (perfect) matching.
    \item Language of planar graphs.
    \item Language of graphs that whichever way their edges are colored using 2 colors, there is a monochromatic triangle.
\end{enumerate}

\lineparagraph {Solution}

a) Using the Witness Theorem (witness for no answer).

According to Hall's marriage theorem there exists a complete matching in a bipartite graph \textbf{if and only if} there is no set of $X$ vertices coming from one side of the bipartite graph for which $|X| > |N(X)|$. (The number of their neighbours is less than their number.)

\begin{itemize}
    \item Witness: A set of $X$ vertices, breaking Hall's condition.
        \begin{itemize}
            \item Since Hall's theorem states that the existence of such a set is equivalent to the bipartite graph having no complete matching, this means that the witness exists for all rejected words, but no accepted words.
            \item The size of the witness is at most $O(n\log_{2}n)$.
        \end{itemize}
    \item Witness checking algorithm:
    \begin{itemize}
        \item Check that all vertices exist, are on the same side of the bipartite graph and count their neigbhours.
        \item It can be shown that this runs in polynomial time.
    \end{itemize}
\end{itemize}

b) Using the Witness Theorem: Similar to the previous one, using Tutte's Theorem.

Tutte's Theorem: A (generic) graph has a perfect matching if and only if for every subset $U$ its vertices, removing that $U$ from the graph: $G \setminus{} U$ falls apart to having at most $|U|$ odd components.

Then, the witness is such an $U$, etc.

c) Using the Witness Theorem: Similar to the previous one, using Kuratowski's Theorem.

Kruskal's Theorem: A graph is planar if and only if it does not contain a subgraph that is a subdivision of $K_5$ or $K_{3,3}$.

Then, the witness is such a subgraph, etc.