\subsection{Session 3, Exercise 5}

\lineparagraph{Exercise}

Give regular expressions for the languages over alphabet $\{0,1\}$ that consist of the following words.

\begin{enumerate}[(a)]
\item Words of odd lengths.
\item Words of even length that start and end with $1$.
\item Words containing at least three $0$'s.
\item Words containing an even number of $0$'s.
\item Words of odd lengths starting with $0$ and words of even length starting with $1$.
\item Words of odd length containing subword $00$.
\end{enumerate}

\lineparagraph{Solution}

\subsubsection{Words of odd lengths}

From the previous exercise we know, that $((0+1)(0+1))^*$ accepts the words of even lengths. If we add a $(0+1)$ at the beginning it will add one more character to the lengths, making them odd: $(0+1)((0+1)(0+1))^*$.

\subsubsection{Words of even length that start and end with 1}

To start and end with $1$'s, the regular expression will be $1\text{<something here>}1$. For <something here>, we need to add a regular expression, that together with the two other $1$'s will allow for an even number of characters. So without the two $1$'s, we need an even number of characters, for which we know the regular expression: $((0+1)(0+1))^*$. Putting these together we arrive at $1((0+1)(0+1))^*1$. Since $((0+1)(0+1))^*$ accepts the empty string, the final regular expression will also accept $11$, which is the shortest possible string in the language.

\subsubsection{Words of odd length that start and end with 1}

This was not in the exercise, however I would like to illustrate a point here.

Let's follow the same pattern of thought, as in the previous exercise:
\begin{itemize}
    \item To start and end with $1$'s, the regular expression will be $1\text{<something here>}1$. \item For <something here>, we need to add a regular expression, that together with the two other $1$'s will allow for an odd number of characters.
    \item So without the two $1$'s, we need an odd number of characters, for which we know the regular expression: $(0+1)((0+1)(0+1))^*$.
    \item Putting these together we arrive at $1(0+1)((0+1)(0+1))^*1$.
    \item We have made a mistake...
\end{itemize}

What is the shortest word in this language? It's $1$, which is not accepted by the regular expression above. The issue is that we did not think about the fact, that both starting and ending in a $1$ can also literally mean the same $1$ character, nothing more.

In general, it is always important when building regular expressions from multiple parts, to check for short strings in the language, to see if our expression works for the simplest cases as well.

In this example, to fix the regular expression above, we simply append the missing case at the end: $1(0+1)((0+1)(0+1))^*1 + 1$. This accepts words of length $1$, $3$, and so on, which is what we've wanted.

\subsubsection{Words containing at least three 0's}

Anything can go between the $0$'s, so simply $(0+1)^*0(0+1)^*0(0+1)^*0(0+1)^*$. The three $0$'s between the $(0+1)^*$'s enforce that the string must contain at least three $0$'s.

\subsubsection{Words containing an even number of 0's}

Let's start by words containing exactly two $0$'s: $1^*01^*01^*$. Then, to allow an even number of $0$'s, we can use the $^*$ operator on this: $(1^*01^*01^*)^*$. However, this regular expression has one problem: if the number of $0$'s is zero, we are unable to match anything other than the empty string. However, we would like to match $1^*$, so we can fix this, by either simply appending it at the end: $(1^*01^*01^*)^* + 1^*$, or to make it shorter, simply moving the first one outside of the outer $^*$ operator: $1^*(01^*01^*)^*$.

\subsubsection{Words of odd lengths starting with 0 and words of even length starting with 1}

Let's make two regular expressions and combine them with a $+$ at the end.

Words of odd lengths starting with $0$: $0((0+1)(0+1))^*$. $0$ matching literal $0$ at the beginning, then $((0+1)(0+1))^*$, matching the remaining even number of any characters, which in total result in an odd number of characters.

Words of even length starting with $1$: $1(0+1)((0+1)(0+1))^*$. Since the word must start with a $1$, this no longer can be an empty string. The shortest possible words are $10$ and $11$, which are correctly matched, and can be followed by an even number of characters.

Finally, combinging the two: $0((0+1)(0+1))^* + 1(0+1)((0+1)(0+1))^*$, or to make it shorter: $(0 + 1(0+1))((0+1)(0+1))^*$.

\subsubsection{Words of odd length containing subword 00}

$00$ is even length, so it either starts with an odd length string, then $00$, then an even length string or vice versa: $((0+1)(0+1))^*(0+1)00((0+1)(0+1))^* + ((0+1)(0+1))^*00(0+1)((0+1)(0+1))^*$ and we can shorten this a little like this: $((0+1)(0+1))^*((0+1)00+00(0+1))((0+1)(0+1))^*$. Again, paying attention to the shortest words in the language, which are of length three, this one can match those as well.

